---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# EPGminer

<!-- badges: start -->
[![R-CMD-check](https://github.com/LylChun/epgminer/workflows/R-CMD-check/badge.svg)](https://github.com/LylChun/epgminer/actions)
<!-- badges: end -->

Electrical Penetration Graph (EPG) is a technique used to monitor insect feeding behavior. Studying voltage waveforms in EPG data allows one to analyze the insect-host relationship and gain insights into what makes a suitable host.

To facilitate the analysis of EPG data, `EPGminer` contains various functions for working with voltage time series data as well as ANA annotation files from manual annotation software. 

The primary utility of using `EPGminer` for analysis is the ability to calculate Frequency and Relative Amplitude. Frequency is calculated using the Fourier Transform to extract the main constituent frequencies in each waveform. Relative amplitude is a measure of the voltage amplitude for each waveform.

Another aspect of `EPGminer` is data visualization via plots. The plotting functions are all in the format plot_x where x denotes the type of plot. The types of plots currently supported are: voltage time series, labeled time-series, Fourier transformed data, frequency boxplot, and two types of pie charts.

An accompanying Shiny app, available as both a web version (https://epgdata.shinyapps.io/epgminer_app/) and locally through R, includes a streamlined version of the package functionality focusing on the main calculations and visuals.

To cite the current version (with DOI):

Elizabeth Chun. (2022). LylChun/EPGminer: EPGminer (v1.0.0). Zenodo. [![DOI](https://zenodo.org/badge/341639621.svg)](https://zenodo.org/badge/latestdoi/341639621)


## System requirements

### Hardware
`EPGminer` can be run on any standard computer with enough memory to perform the necessary operations. It was primarily tested on a computer with 8 GB of RAM which is amply sufficient.

### Software
#### OS
The package was developed and primarily tested on Windows 10 Version 20H2. It has also been tested on macOS **Insert Here**

#### Dependencies
This R package was developed using R version 4.1.0. It depends on R (>=4.0.0) as well as specific package Imports listed in the DESCRIPTION file.

#### EPGminer Application
A streamlined version of `EPGminer` is available as a Shiny application. The `EPGminer` App web version may be launched via the user's web browser with no installation or other software dependencies required. The app has been tested in Chrome, Edge, and **insert**. Note that due to server limitations, the web version is limited in memory and should not be used for datasets that exceed 100mb at most.

The `EPGminer` app local version must be installed with the R package itself and System Requirements are as noted above.

## Installation guide
To use the `EPGminer` package, users must have R installed on their computer. This may be accomplished by going to https://cran.r-project.org/, downloading, and installing the appropriate version of R for the given operating system. Installation of R should take around 1 minute.

Once R has been properly installed, the `EPGminer` package may be installed from GitHub as follows.

### Install from GitHub
The `EPGminer` package can be installed from GitHub using `devtools` via the following code in R. Installation should take around 20-30 seconds.

``` {r, eval = FALSE}
if (!require(devtools)){
  install.packages("devtools")
}
devtools::install_github("LylChun/EPGminer", build_vignettes = TRUE)
```

## Demo
For a demo of `EPGminer` with instructions, expected output, and expected runtime, please see the "EPGminerDemo" vignette (to be filled in):

```{r, eval = F}
vignette("EPGminerDemo")
```

### Instructions for use
Use of the software for EPG data analysis involves data reading/processing, labeling, and calculation of metrics. For an example script showing how we used `EPGminer` on our data, please see the "EPGminerReprex" vignette:

```{r, eval = F}
vignette("EPGminerReprex")
```

